---
title: "The Illinois People Over Parking Act"
author: "Michael McLean - Member, Abundant Housing Illinois"
date: "March 14, 2025"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide
    self_contained: true
    output_dir: "docs"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction

### Understanding Parking Mandates

Parking mandates, also known as parking minimums, are zoning regulations that require developers to build a minimum number of parking spaces with new housing and commercial developments. These mandates typically specify a certain number of parking spaces per unit of housing, square footage of retail space, or other metrics depending on the development type.

#### Why Parking Mandates Are Costly:

-   **Increased Housing Costs**: Each parking space can cost \$30,000-\$75,000 to build, significantly raising the cost of housing development and ultimately housing prices.
-   **Reduced Housing Supply**: Land used for parking cannot be used for additional housing units, limiting the overall housing supply.
-   **Environmental Impact**: Excessive parking encourages car dependency, increasing traffic congestion and carbon emissions.
-   **Inefficient Land Use**: Parking lots create "dead spaces" in urban areas that could otherwise be used for housing, businesses, or public spaces.
-   **Economic Burden**: Many parking spaces sit empty much of the time, representing wasted resources and opportunity costs.

#### Why Parking Mandates Should Be Repealed:

-   **Market-Based Solutions**: Developers can better determine the appropriate amount of parking based on actual demand rather than arbitrary requirements.
-   **Transit-Oriented Development**: Eliminating parking mandates near transit encourages development that leverages existing public transportation infrastructure.
-   **Affordability**: Reducing or eliminating parking requirements can make housing more affordable and accessible.
-   **Sustainability**: Less parking promotes walking, cycling, and public transit use, reducing carbon emissions.
-   **Vibrant Communities**: Space previously dedicated to parking can be repurposed for housing, businesses, and community amenities.

### Enter: The People Over Parking Act
Let's analyzes and visualize the potential impact of the Illinois [People Over Parking Act (HB3256)](https://www.ilga.gov/legislation/billstatus.asp?DocNum=3256&GAID=18&GA=104&DocTypeID=HB&LegID=161742&SessionID=114) on Chicago. The proposed legislation prohibits local governments from imposing minimum parking requirements on development projects located within 1/2 mile of public transportation hubs.

According to the bill, a "public transportation hub" is defined as:

-   A rail transit station
-   A boat or ferry terminal served by either a bus connection stop or rail transit station
-   A bus connection stop of 2 or more major bus routes with a frequency of service interval of 15 minutes or less during peak commute periods

Where will land qualify under those parameters?

```{r packages}
# Install packages if not already installed
required_packages <- c("tidyverse", "sf", "leaflet", "leaflet.extras", 
                       "data.table", "zip", "httr", "lubridate", "mapview")

new_packages <- required_packages[!required_packages %in% installed.packages()[,"Package"]]
if(length(new_packages)) install.packages(new_packages)

# Load required packages
library(tidyverse)
library(sf)
library(leaflet)
library(leaflet.extras)
library(data.table)
library(zip)
library(httr)
library(lubridate)
library(mapview)
```


```{r download_gtfs, results='hide'}
## Download and Process GTFS Data

## We'll download the latest GTFS data for all three transit agencies (CTA, Pace, and Metra),
## extract them, and process them to identify public transportation hubs according to the bill's definition.

# Function to download and extract GTFS data
download_and_extract_gtfs <- function(agency_name, zip_link) {
  # Create a temporary directory to store the downloaded file
  temp_dir <- file.path(tempdir(), agency_name)
  if (!dir.exists(temp_dir)) {
    dir.create(temp_dir, recursive = TRUE)
  }
  
  temp_file <- file.path(temp_dir, paste0(agency_name, "_gtfs.zip"))
  
  # Create a cache directory
  cache_dir <- "gtfs_cache"
  if (!dir.exists(cache_dir)) {
    dir.create(cache_dir)
  }
  cache_file <- file.path(cache_dir, paste0(agency_name, "_gtfs.zip"))
  
  # Try to download or use cached data
  tryCatch({
    # Download the GTFS ZIP file with a timeout and user agent
    options(timeout = 60)  # Increase timeout to 60 seconds
    
    # Use httr::GET with a user agent to avoid 403 Forbidden errors
    response <- httr::GET(
      zip_link,
      httr::user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/91.0.4472.124 Safari/537.36"),
      httr::write_disk(temp_file, overwrite = TRUE),
      httr::timeout(60)
    )
    
    # Check if the request was successful
    if (httr::status_code(response) != 200) {
      stop(paste0("Failed to download with status code: ", httr::status_code(response)))
    }
    
    # Save a copy to cache
    file.copy(temp_file, cache_file, overwrite = TRUE)
    
  }, error = function(e) {
    message(paste0("Download failed for ", agency_name, ": ", e$message))
    if (file.exists(cache_file)) {
      message(paste0("Using cached GTFS data for ", agency_name, " from ", cache_file))
      file.copy(cache_file, temp_file, overwrite = TRUE)
    } else {
      stop(paste0("Could not download GTFS data for ", agency_name, " and no cache available."), call. = FALSE)
    }
  })
  
  # Extract the files
  gtfs_files <- unzip(temp_file, exdir = temp_dir)
  
  # Return the directory containing the extracted files
  return(temp_dir)
}

# Download and extract GTFS data for all three agencies
cta_dir <- download_and_extract_gtfs("cta", "https://www.transitchicago.com/downloads/sch_data/google_transit.zip")
pace_dir <- download_and_extract_gtfs("pace", "https://www.pacebus.com/sites/default/files/2025-02/GTFS.zip")
metra_dir <- download_and_extract_gtfs("metra", "https://schedules.metrarail.com/gtfs/schedule.zip")
```
```{r process_stops}
## Identify Public Transportation Hubs

## Now we'll process the GTFS data from all three agencies to identify the public transportation hubs as defined in the legislation.

# Function to read and normalize GTFS data
read_normalize_gtfs <- function(agency_name, agency_dir) {
  # Read the stops data
  stops_file <- file.path(agency_dir, "stops.txt")
  if (file.exists(stops_file)) {
    stops <- fread(stops_file)
    # Add agency identifier
    stops[, agency := agency_name]
    
    # Normalize column names and types
    if (!"location_type" %in% names(stops)) {
      stops[, location_type := NA_integer_]
    }
    if (!"parent_station" %in% names(stops)) {
      stops[, parent_station := NA_character_]
    }
    
    # Ensure stop_id is character
    stops[, stop_id := as.character(stop_id)]
    
    # Create a unique ID that includes the agency
    stops[, unique_stop_id := paste0(agency_name, "_", stop_id)]
  } else {
    stops <- data.table(
      stop_id = character(),
      stop_name = character(),
      stop_lat = numeric(),
      stop_lon = numeric(),
      location_type = integer(),
      parent_station = character(),
      agency = character(),
      unique_stop_id = character()
    )
  }
  
  # Read the routes data
  routes_file <- file.path(agency_dir, "routes.txt")
  if (file.exists(routes_file)) {
    routes <- fread(routes_file)
    # Add agency identifier
    routes[, agency := agency_name]
    
    # Normalize route_id to character
    routes[, route_id := as.character(route_id)]
    
    # Create a unique ID that includes the agency
    routes[, unique_route_id := paste0(agency_name, "_", route_id)]
  } else {
    routes <- data.table(
      route_id = character(),
      route_type = integer(),
      agency = character(),
      unique_route_id = character()
    )
  }
  
  # Read trips data
  trips_file <- file.path(agency_dir, "trips.txt")
  if (file.exists(trips_file)) {
    trips <- fread(trips_file)
    # Add agency identifier
    trips[, agency := agency_name]
    
    # Normalize trip_id and route_id to character
    trips[, trip_id := as.character(trip_id)]
    trips[, route_id := as.character(route_id)]
    
    # Create unique IDs that include the agency
    trips[, unique_trip_id := paste0(agency_name, "_", trip_id)]
    trips[, unique_route_id := paste0(agency_name, "_", route_id)]
  } else {
    trips <- data.table(
      trip_id = character(),
      route_id = character(),
      service_id = character(),
      agency = character(),
      unique_trip_id = character(),
      unique_route_id = character()
    )
  }
  
  # Read stop_times data
  stop_times_file <- file.path(agency_dir, "stop_times.txt")
  if (file.exists(stop_times_file)) {
    stop_times <- fread(stop_times_file)
    # Add agency identifier
    stop_times[, agency := agency_name]
    
    # Normalize trip_id and stop_id to character
    stop_times[, trip_id := as.character(trip_id)]
    stop_times[, stop_id := as.character(stop_id)]
    
    # Create unique IDs that include the agency
    stop_times[, unique_trip_id := paste0(agency_name, "_", trip_id)]
    stop_times[, unique_stop_id := paste0(agency_name, "_", stop_id)]
  } else {
    stop_times <- data.table(
      trip_id = character(),
      stop_id = character(),
      arrival_time = character(),
      departure_time = character(),
      stop_sequence = integer(),
      agency = character(),
      unique_trip_id = character(),
      unique_stop_id = character()
    )
  }
  
  # Read calendar data
  calendar_file <- file.path(agency_dir, "calendar.txt")
  if (file.exists(calendar_file)) {
    calendar <- fread(calendar_file)
    # Add agency identifier
    calendar[, agency := agency_name]
  } else {
    calendar <- data.table(
      service_id = character(),
      monday = integer(),
      tuesday = integer(),
      wednesday = integer(),
      thursday = integer(),
      friday = integer(),
      saturday = integer(),
      sunday = integer(),
      start_date = integer(),
      end_date = integer(),
      agency = character()
    )
  }
  
  # Return all normalized tables
  return(list(
    stops = stops,
    routes = routes,
    trips = trips,
    stop_times = stop_times,
    calendar = calendar
  ))
}

# Read and normalize GTFS data for all three agencies
cta_data <- read_normalize_gtfs("cta", cta_dir)
pace_data <- read_normalize_gtfs("pace", pace_dir)
metra_data <- read_normalize_gtfs("metra", metra_dir)

# Combine data from all agencies
all_stops <- rbindlist(list(cta_data$stops, pace_data$stops, metra_data$stops), fill = TRUE)
all_routes <- rbindlist(list(cta_data$routes, pace_data$routes, metra_data$routes), fill = TRUE)
all_trips <- rbindlist(list(cta_data$trips, pace_data$trips, metra_data$trips), fill = TRUE)
all_stop_times <- rbindlist(list(cta_data$stop_times, pace_data$stop_times, metra_data$stop_times), fill = TRUE)
all_calendar <- rbindlist(list(cta_data$calendar, pace_data$calendar, metra_data$calendar), fill = TRUE)

# Identify rail transit stations across all agencies
# CTA: route_type = 1 (subway/metro)
# Metra: route_type = 2 (rail)
# Pace: No rail routes
rail_routes <- all_routes[route_type %in% c(1, 2)]

# For CTA, use parent_station or location_type to identify stations
cta_rail_stops <- all_stops[
  agency == "cta" & 
  ((!is.na(parent_station) & parent_station != "") | 
   (!is.na(location_type) & location_type == 1))
]

# For Metra, all stops are rail stations, but filter out Wisconsin stations
# The Illinois-Wisconsin border is approximately at 42.5 degrees latitude
metra_rail_stops <- all_stops[agency == "metra" & stop_lat <= 42.5]

# Combine all rail stations
rail_stops <- rbindlist(list(cta_rail_stops, metra_rail_stops), fill = TRUE)

# Create a spatial object for rail stations
rail_stations_sf <- st_as_sf(rail_stops, coords = c("stop_lon", "stop_lat"), crs = 4326)

# Now identify bus stops that meet the criteria of 2+ major routes with frequency ≤ 15 min
# First determine the service periods for weekdays (Monday-Friday)
weekday_service <- all_calendar[
  monday == 1 & tuesday == 1 & wednesday == 1 & thursday == 1 & friday == 1, 
  .(service_id, agency)
]

# Get the trips that operate on weekdays
weekday_trips <- merge(all_trips, weekday_service, by = c("service_id", "agency"))

# Define peak hours (e.g., 7-9 AM and 4-6 PM)
morning_peak_start <- as.ITime("07:00:00")
morning_peak_end <- as.ITime("09:00:00")
evening_peak_start <- as.ITime("16:00:00")
evening_peak_end <- as.ITime("18:00:00")

# Process stop times for peak hours
all_stop_times[, arrival_time_hhmmss := substr(arrival_time, 1, 8)]
all_stop_times[, arrival_time_obj := as.ITime(arrival_time_hhmmss)]

peak_stop_times <- all_stop_times[
  (arrival_time_obj >= morning_peak_start & arrival_time_obj <= morning_peak_end) |
  (arrival_time_obj >= evening_peak_start & arrival_time_obj <= evening_peak_end)
]

# Join with trips to get route information
peak_stop_times <- merge(
  peak_stop_times, 
  weekday_trips[, .(unique_trip_id, unique_route_id, agency)], 
  by = c("unique_trip_id", "agency")
)

# Count unique routes per stop during peak hours
routes_per_stop <- peak_stop_times[, .(unique_routes = uniqueN(unique_route_id)), by = .(unique_stop_id, agency)]

# Get stops with 2+ routes
multi_route_stops <- routes_per_stop[unique_routes >= 2]

# Filter to bus stops only (not rail stations)
# For CTA: location_type == 0 or NA, and not in rail_stops
# For Pace: All stops are bus stops
# For Metra: None (all are rail)
bus_stops <- all_stops[
  (agency == "cta" & (is.na(location_type) | location_type == 0) & 
   !(unique_stop_id %in% rail_stops$unique_stop_id)) |
  (agency == "pace")
]

# Calculate the frequency for each route at each stop during peak hours
# First, separate morning and evening peak periods
morning_peak_times <- peak_stop_times[
  arrival_time_obj >= morning_peak_start & arrival_time_obj <= morning_peak_end
]

evening_peak_times <- peak_stop_times[
  arrival_time_obj >= evening_peak_start & arrival_time_obj <= evening_peak_end
]

# Function to calculate headways for a given set of stop times
calculate_headways <- function(stop_times_data) {
  # Sort by stop, route, and time
  setorder(stop_times_data, unique_stop_id, unique_route_id, arrival_time_obj)
  
  # Calculate time difference between consecutive arrivals of the same route at the same stop
  stop_times_data[, time_diff := c(NA, diff(as.numeric(arrival_time_obj))), 
                 by = .(unique_stop_id, unique_route_id)]
  
  # Convert time difference from seconds to minutes
  stop_times_data[, headway_minutes := time_diff / 60]
  
  # Filter out unreasonable headways (e.g., when there's a large gap between trips)
  stop_times_data <- stop_times_data[!is.na(headway_minutes) & headway_minutes <= 60]
  
  return(stop_times_data)
}

# Calculate headways for morning and evening peak periods
morning_headways <- calculate_headways(morning_peak_times)
evening_headways <- calculate_headways(evening_peak_times)

# Combine morning and evening headways
all_headways <- rbind(morning_headways, evening_headways)

# Calculate median headway for each route at each stop
route_headways <- all_headways[, .(
  median_headway = median(headway_minutes, na.rm = TRUE),
  min_headway = min(headway_minutes, na.rm = TRUE),
  max_headway = max(headway_minutes, na.rm = TRUE),
  num_observations = .N
), by = .(unique_stop_id, unique_route_id, agency)]

# Filter to routes with sufficient observations (at least 3)
route_headways <- route_headways[num_observations >= 3]

# Identify routes that meet the 15-minute frequency criterion
route_headways[, meets_frequency := median_headway <= 15]

# For each stop with multiple routes, check if all routes meet the frequency criterion
stop_route_counts <- route_headways[, .(
  total_routes = .N,
  qualifying_routes = sum(meets_frequency),
  all_routes_qualify = all(meets_frequency)
), by = .(unique_stop_id, agency)]

# Filter to stops with 2+ routes
multi_route_stops_with_frequency <- stop_route_counts[total_routes >= 2]

# Identify stops where all routes meet the frequency criterion
qualifying_stops <- multi_route_stops_with_frequency[all_routes_qualify == TRUE]

# Get the full stop information for qualifying stops
qualifying_bus_hubs <- merge(
  bus_stops,
  qualifying_stops[, .(unique_stop_id, agency, total_routes, qualifying_routes)],
  by = c("unique_stop_id", "agency")
)

# Find bus stops that meet both criteria: multiple routes and frequency requirement
bus_hub_candidates <- qualifying_bus_hubs

# Create a spatial object for bus hubs
bus_hubs_sf <- st_as_sf(bus_hub_candidates, coords = c("stop_lon", "stop_lat"), crs = 4326)

# Ensure both spatial objects have the same columns before combining
# First, identify common columns
rail_cols <- names(rail_stations_sf)
bus_cols <- names(bus_hubs_sf)

# Add missing columns to each dataset
for (col in setdiff(bus_cols, rail_cols)) {
  rail_stations_sf[[col]] <- NA
}
for (col in setdiff(rail_cols, bus_cols)) {
  bus_hubs_sf[[col]] <- NA
}

# Add type column to both
rail_stations_sf$type <- "rail"
bus_hubs_sf$type <- "bus_hub"

# Now combine them
all_hubs_sf <- rbind(rail_stations_sf, bus_hubs_sf)

# Add agency information to the combined hubs
all_hubs_sf$agency_name <- factor(
  all_hubs_sf$agency,
  levels = c("cta", "pace", "metra"),
  labels = c("CTA", "Pace", "Metra")
)
```
```{r create_buffers}
## Create 1/2 Mile Buffers Around Transportation Hubs
## Now we'll create a 1/2 mile buffer around each transportation hub to visualize the areas affected by the legislation.

# Convert to a projected CRS for accurate buffer calculation
# NAD83 / Illinois East (ftUS) EPSG:3435 is appropriate for Chicago
all_hubs_projected <- st_transform(all_hubs_sf, 3435)

# Create a 1/2 mile buffer (2640 feet)
half_mile_buffers <- st_buffer(all_hubs_projected, 2640)

# Union all buffers to create a single polygon that shows all affected areas
all_affected_areas <- st_union(half_mile_buffers)

# Convert back to WGS84 for mapping
all_affected_areas_wgs84 <- st_transform(all_affected_areas, 4326)
half_mile_buffers_wgs84 <- st_transform(half_mile_buffers, 4326)
```

## Where would you be able to build a home without required parking?


```{r create_map}

# We'll use Leaflet to create an interactive map showing the affected areas.

# Define a color palette for transit agencies
agency_pal <- colorFactor(
  palette = c("#009CDE", "#814C9E", "#E31837"),  # CTA blue, Pace purple, Metra red
  domain = all_hubs_sf$agency_name
)

# Define a color palette for transit hub types
type_pal <- colorFactor(
  palette = c("blue", "green"),
  domain = all_hubs_sf$type
)

# Create a leaflet map
map <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = all_affected_areas_wgs84,
    fillColor = "purple",
    fillOpacity = 0.25,
    weight = 1,
    color = "purple",
    opacity = 0.7,
    group = "Affected Areas (1/2 Mile from Hubs)"
  ) %>%
  # Add points for transit hubs, colored by agency
  addCircleMarkers(
    data = all_hubs_sf,
    radius = 3,
    color = ~agency_pal(agency_name),
    stroke = FALSE,
    fillOpacity = 0.8,
    group = "Transit Hubs by Agency",
    popup = ~paste0(
      "<strong>", stop_name, "</strong><br>",
      "Agency: ", agency_name, "<br>",
      "Type: ", type, "<br>",
      "Stop ID: ", stop_id
    )
  ) %>%
  # Add layer controls
  addLayersControl(
    baseGroups = c("CartoDB Positron"),
    overlayGroups = c("Affected Areas (1/2 Mile from Hubs)", "Transit Hubs by Agency"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  # Add legend for agencies
  addLegend(
    position = "bottomright",
    colors = c("purple", "#009CDE", "#814C9E", "#E31837"),
    labels = c("Transit-served Areas Affected by the People Over Parking Act", 
               "CTA", "Pace", "Metra"),
    opacity = 0.7
  ) %>%
  addFullscreenControl() %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "miles",
    primaryAreaUnit = "sqmiles",
    activeColor = "#3D535D",
    completedColor = "#7D4479"
  ) %>%
  addMiniMap(
    tiles = providers$CartoDB.Positron,
    toggleDisplay = TRUE
  )

# Display the map
map
```

## How much of Chicago would be free from parking mandates?

```{r analyze_areas}
# Calculate the total area affected in square miles
affected_area_sqft <- st_area(all_affected_areas)
affected_area_sqmi <- units::set_units(affected_area_sqft, "mi^2")

# Get a rough estimate of Chicago's total area (approximately 234 sq mi)
chicago_area_sqmi <- 234

# Calculate percentage of Chicago potentially affected
pct_affected <- as.numeric(affected_area_sqmi) / chicago_area_sqmi * 100

# Count the number of transit hubs by agency
hub_counts <- table(all_hubs_sf$agency_name)

# For a more accurate analysis, we would use actual city boundaries
# and calculate the exact intersection, but this gives a rough estimate
```

The analysis shows that approximately `r round(as.numeric(affected_area_sqmi), 2)` square miles of Chicago's area would be affected by the People Over Parking Act, which is roughly `r round(pct_affected, 1)`% of the city's total area.

### Note on Bus Transit Hub Criteria

This analysis strictly applies the frequency criterion from the legislation, requiring all bus routes at a hub to have service intervals of 15 minutes or less during peak periods. Our verification shows:

- CTA routes: 99.98% meet the 15-minute frequency criterion
- Pace routes: Only 16.04% meet the 15-minute frequency criterion

This difference in service frequency significantly impacts which areas qualify under the legislation, with a much higher concentration of qualifying hubs in areas served by CTA (primarily Chicago) compared to suburban areas served by Pace.

```{r agency_comparison, fig.width=8, fig.height=4}
# Create a data frame for visualization
agency_stats <- route_headways[, .(
  total_routes = .N,
  qualifying_routes = sum(meets_frequency),
  pct_qualifying = sum(meets_frequency) / .N * 100
), by = agency]

# Create a bar chart showing the percentage of routes meeting the frequency criterion by agency
ggplot(agency_stats, aes(x = agency, y = pct_qualifying, fill = agency)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(pct_qualifying, 1), "%")), 
            position = position_stack(vjust = 0.5), color = "white", size = 5) +
  scale_fill_manual(values = c("cta" = "#009CDE", "pace" = "#814C9E")) +
  labs(
    title = "Percentage of Routes Meeting 15-Minute Frequency Criterion",
    subtitle = "By Transit Agency",
    x = "Agency",
    y = "Percentage of Routes",
    fill = "Agency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```

### Transit Hub Distribution by Agency

The analysis includes transit hubs from three agencies:

- CTA: `r hub_counts["CTA"]` transit hubs
- Pace: `r hub_counts["Pace"]` transit hubs
- Metra: `r hub_counts["Metra"]` transit hubs

This comprehensive approach ensures that all qualifying transit hubs in the Chicagoland region are included in the analysis, providing a more accurate representation of areas that would be affected by the People Over Parking Act.

## Resources and References

-   [People Over Parking Act (HB3256)](https://www.ilga.gov/legislation/billstatus.asp?DocNum=3256&GAID=18&GA=104&DocTypeID=HB&LegID=161742&SessionID=114) [SB2352](https://ilga.gov/legislation/BillStatus.asp?DocNum=2352&GAID=18&DocTypeID=SB&LegId=162316&SessionID=114&GA=104)
-   [Chicago Transit Authority GTFS Data](https://www.transitchicago.com/downloads/sch_data/)
-   [General Transit Feed Specification Reference](https://developers.google.com/transit/gtfs/reference)
-   [Illinois Homes For All Coalition](https://ilhomesforall.org/)
-   [Abundant Housing Illinois](https://abundanthousingillinois.org/)
-   [IPM News - "Housing costs in Illinois are rising. Lawmakers are considering several bills that could help"](https://ipmnewsroom.org/housing-costs-in-illinois-are-rising-lawmakers-are-considering-several-bills-that-could-help/)
-   [Hyde Park Herald "Buckner-sponsored bills aim to speed up housing development in Illinois cities"](https://www.hpherald.com/evening_digest/buckner-sponsored-bills-aim-to-speed-up-housing-development-in-illinois-cities/article_3e6d29a2-f623-11ef-9007-b74a49539892.html)
-   [Evanston Round Table "New zoning bills would enact some Envision Evanston proposals statewide"](https://evanstonroundtable.com/2025/02/03/new-bills-would-take-some-envision-zoning-proposals-statewide/)
-   [Illinois REALTORS unveils legislative agenda aimed at easing housing shortage](https://abundanthousingillinois.org/op-eds/illinois-realtors-unveils-legislative-agenda-aimed-at-easing-housing-shortage/)
-   [Parking Reform Network "What is parking reform?"](https://parkingreform.org/what-is-parking-reform/)
